GPU-friendly Stroke Expansion

### 2024-07-28 paper.tex

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves (also known as ``offset curves'') of path segments, combined with caps and the outer contours of joins. Both \citet{Nehab2020} and \citet{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, \emph{evolute} segments are required in addition to the parallel curves, as well as \emph{inner contours of joins} when segments are short. The technique described in this paper produces strongly correct outlines according to this definition. Examples of weakly and strongly correct rendering are shown in Figure~\ref{fig:weak_strong}.

@@

The subproblem of approximating parallel curves has spawned extensive literature, none of which is entirely satisfying, especially when it comes to algorithms that can be efficiently evaluated on GPU. An example of a reasonably good algorithm that computes flattened parallel curves is \citet{Yzerman2020}, as used in the Blend2D rendering library. For producing curved outlines, the Tiller and Hanson~\cite{Tiller1984} algorithm is commonly implemented and cited, but its performance is quite poor when applied to cubic Béziers. The quadratic Bézier version is adequate, and forms the basis of the algorithm in \citet{Nehab2020}. A variant is also used in Skia~\cite{Skia}; instead of lowering the cubic Bézier to a quadratic approximation and then computing the parallel curve, Skia approximates the parallel curve directly with a quadratic Bézier, then measures the error to determine whether further subdivision is necessary.

Previous work on stroke rendering does not fully solve the problem of fast and correct stroking. Most existing implementations have correctness problems, in particular failing to draw the evolutes and inner joins as required for strong correctness; \citet{Nehab2020} provides a detailed survey. In addition, most implementations are not suitable for running on GPU, requiring sequential execution to produce watertight outlines. One GPU implementation is polar stroking (\citet{Kilgard2020}), a local technique, but among other limitations, it does not bound the Fréchet distance error (rather, it bounds the angle error).

@@

Previous work, particularly \citet{Kilgard2020}, has explained how to break the input path into dash segments of the specified length on GPU using a prefix sum of segment arc lengths. There are implementation tradeoffs to GPU-based dash processing, including potentially a larger number of dispatch stages. We anticipate that the potential performance benefit of implementing dashing on a GPU depends on the scene and other factors in the overall rendering pipeline. While we would expect further performance improvement from GPU-based dashing, we did not fully explore the solution space and we were able to achieve acceptable performance with dash processing on CPU.

@@

There are generally three approaches to such curve approximation. The most straightforward but also least efficient is ``cut then measure,'' usually combined with adaptive subdivision. In this technique, a candidate approximate curve is produced, then the error is measured against the source curve, usually by sampling a number of points along both curves and determining a maximum error (or perhaps some other error norm). If the error is within tolerance, the approximation is accepted. Otherwise, the curve is subdivided (usually at $t = 0.5$) and each half is recursively approximated. A substantial fraction of all curve approximation methods in the literature are of this form, including \citet{Nehab2020}. The main disadvantage is the cost of computing the error metric. Another risk is underestimating the error due to inadequate sampling; this is a particular problem when the source curve contains a cusp.

@@

By far the most efficient approach is an \emph{invertible} error metric. In this approach, the error metric has an analytic inverse, or at least a good numerical approximation. Because the metric is invertible, it can predict the number of subdivisions needed, as well as the parameter value for each subdivision. If the error metric is accurate, then approximation is near-optimal. One example of an invertible error metric is angle step, used in polar stroking (\citet{Kilgard2020}); the number of subdivisions is the total angle subsumed by the curve divided by the angle step size, and the parameter value for each subdivision is the result of solving for a tangent direction. Another widely used invertible error metric is Wang's formula (\citet{Goldman2003}, Section 5.6.3), which gives a bound on the flattening error based on the second derivative of the curve. This metric is conservative but works well in practice; among other applications, it is used in Skia for path flattening. The chief limitation of Wang's method is that it computes a subdivision bound for a Bézier curve without accounting for the displacement of the parallel curve due to stroking. When applied naively to generation of parallel curves, it can undershoot substantially, especially near cusps.

@@

It is common to approximate cubic Béziers to some intermediate curve format more conducive to offsetting and flattening. A number of published solutions (\citet{Yzerman2020}, \citet{Nehab2020}) use quadratic Béziers, as it is well suited for computation of parallel curves. Even so, this curve has some disadvantages. For one, it cannot model an inflection point, so the source curve must be subdivided at inflection points.

@@

The parallel curve of the Euler spiral (also known as ``clothoid'') was characterized by \citet{Wieleitner1907} well over a hundred years ago, and has a straightforward closed-form Cesàro representation, curvature as a function of arc length, of the form $\kappa(s) = c_0(s - s_0)^{-1/2} + c_1$. Euler spirals as an intermediate curve representation for computing parallel curves was proposed in \citet{Levien2021}, which also expands on the Wieleitner reference (including an English translation) and Cesàro equation.

@@

The problem of approximating a curve by a sequence of arc segments has extensive literature, but none of the published solutions are quite suitable for our application. The specific problem of approximating an Euler spiral by arcs is considered in \citet{Meek2004} using a ``cut then measure'' adaptive subdivision scheme, but their solution has poor quality; it scales as $O(1/n^2)$, while $O(1/n^3)$ is attainable. The result was improved ``slightly'' by \citet{Narayan2014}. The literature also contains optimal results, namely \citet{Maier2014} and \citet{Nuntawisuttiwong2021}, but at considerable cost; both approaches claim $O(n^2)$ time complexity. The through-line for all these results is that they are solving a harder problem: adopting the constraint that the generated arc sequence is $G^1$ continuous. While desirable for many applications, this constraint is not needed for rendering a stroke outline. Even with this constraint relaxed, the angle discontinuities of an arc approximation are tiny compared to flattening to lines.

@@

Many techniques for stroke expansion described in the literature apply some lowering of cubic Bézier curves to a simpler curve type that is more tractable for evaluating the parallel curve. Computing parallel curves directly on cubic Bézier curve segments is not very tractable. In particular, the widely cited Tiller-Hanson algorithm~\cite{Tiller1984} performs well for quadratic Béziers but significantly worse for cubics.

@@

A typical pattern for converting from one curve type to another is \emph{adaptive subdivision.} An approximate curve is found in the parameter space of the target curve family. The error of the approximation is measured. If the error exceeds the specified tolerance, the curve is subdivided (typically at $t = 0.5$), otherwise the approximation is accepted. Subdivisions are also indicated at special points; for example, since quadratic Béziers cannot represent inflection points, and geometric Hermite interpolation is numerically unstable if the input curve is not convex, lowering to quadratic Béziers also requires calculation of inflection points and subdividing there. A good example of this pattern is \citet{Nehab2020}. One advantage of Euler spirals over quadratic Béziers is that they can represent inflection points just fine, so it is not necessary to solve for the inflection points, or do additional subdivision. Avoiding these alleviates the need for conditional logic and case analysis, which makes the algorithm more amenable to GPU execution.

@@

A key step in approximating one curve with another is evaluating the error of the approximation. A common approach (used in \citet{Nehab2020} among others) is to generate the approximate curve, then measure the distance, often by sampling at multiple points on the curve. All this is potentially slow, with the additional risk of underestimating the error due to undersampling.

@@

Given tangent angles relative to the chord, finding the Euler spiral segment that minimizes total curvature variation is a form of geometric Hermite interpolation. There are a number of published solutions to this problem, involving nontrivial numerical solving techniques: gradient descent~\cite{Kimia2003}, bisection~\cite{Walton2009}, or Newton iteration~\cite{Connor2014}. A more direct approach is to approximate the function as a reasonably low-order polynomial in terms of the endpoint angles. Our approach is to use a 7th order polynomial, which is more precise than the 3rd order polynomial proposed in \citet{Reif2021}, at only modest increased cost. The exact coefficients used are presented in Appendix \ref{appendix:gh}.

@@

\begin{figure*}
    \includegraphics[scale=0.15]{stroke_stills}
    \caption{Stroked cubic Béziers exhibiting strong and weak correctness, rendered by our implementation. The contents of the line soup buffer are visualized as green and blue outlines. Segments with a winding direction that points \emph{up} are colored blue while those pointing \emph{down} are colored green. The curves were adapted from the \emph{stills} dataset by \citet{Nehab2020} and the Skia test corpus.}
    \Description{Selection of stroked cubic Béziers exhibiting strong and weak correctness}
    \label{fig:stroke_stills}
\end{figure*}

As recommended in both \citet{Nehab2020} and \citet{Kilgard2020}, the rendering of a cusp with infinite curvature matches that of a near-cusp. The code to detect near-zero derivatives and re-evaluate is a small amount of non-divergent logic, unlike the additional ``regularization'' pass proposed in Section 3.1 of \citet{Nehab2020} to handle special cases.

@@

Even when the input curve is smooth, its parallel curve contains a cusp when the (signed) curvature equals the offset distance (the half-width of the stroke). In published techniques, dealing with these cusps is a nontrivial effort, and involves numerical methods that are not GPU friendly. In particular, detecting locations in the cubic Bézier where the curvature crosses a given quantity is a medium-degree polynomial, and in general requires numerical techniques for root finding. In the approach of \citet{Nehab2020}, the root finder not only requires a hybrid Newton/bisection method, which requires iteration, but is also recursive in that it uses the roots of a polynomial of one degree lower as a subroutine. In general, polynomials up to cubic can be considered GPU-friendly, while finding cusps in cubic Béziers requires a higher degree.

@@

We present two versions of our stroking method: a sequential CPU stroker and the GPU compute shader outlined in Section \ref{section:gpu-impl}. Both versions can generate stroked outlines with line and arc primitives. We focused our evaluation on the number of output primitives and execution time, using the \emph{timings} dataset provided by \citet{Nehab2020}, in addition to our own curve-intensive scenes to stress the GPU implementation. We present our findings in the remainder of this section.

@@

We measured the timing of our CPU implementation, generating both line and arc primitives, against the Nehab and Skia strokers, which both output a mix of straight lines and quadratic Bézier primitives. The results are shown in Figure~\ref{fig:timings}. The time shown is the total for test files from the \emph{timings} dataset of \citet{Nehab2020}, and the tolerance was 0.25 when adjustable as a parameter.

Similarly to \citet{Nehab2020}'s observations, Skia is the fastest stroke expansion algorithm we measured. The speed is attained with some compromises, in particular an imprecise error estimation that can sometimes generate outlines exceeding the given tolerance. We confirmed that this behavior is still present. The Nehab paper proposes a more careful error metric, but a significant fraction of the total computation time is dedicated to evaluating it.

@@

We instrumented our pipeline to obtain fine-grained pipeline stage execution times using the GPU \emph{timestamp query} feature supported by both Metal and Vulkan. This allowed us to collect measurements for the curve flattening and input processing stages in isolation. We ran the compute pipeline on the \emph{timings} SVG files provided by \citet{Nehab2020}. The largest of these SVGs is \emph{waves} (see rendering \emph{(a)} in Figure \ref{fig:renderings}) which contains 42 stroked paths and a total of 13,308 path segments.

@@

\begin{figure}
    \includegraphics[scale=0.69]{nehab_gpu_timings_strong}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the \citet{Nehab2020} \emph{timings} data set. The timings are shown in milliseconds. The hatched bars show the timings for the version that outputs line primitives while the solid bars show that for arcs. Mali-G78 (the lowest-end GPU we tested) converts \emph{waves} (the largest scene in the set) to lines in \textasciitilde3.5 ms. The compute shader executes at least an order of magnitude faster on the other GPUs.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from \citet{Nehab2020} timings SVGs}
    \label{fig:nehab-gpu-timings}
\end{figure}

@@

Figures~\ref{fig:nehab-gpu-timings} and \ref{fig:test-scenes-gpu-timings} show the execution times for the two data sets. The entire \citet{Nehab2020} timings set adds up to less than 1.5~ms on all GPUs except the mobile unit. Our algorithm is \textasciitilde14$\times$ slower on the Mali-G78 compared to the M1 Max but it is still capable of processing \emph{waves} in 3.48~ms on average. We observed that the performance of the kernel scales well with increasing GPU ability even at very large workloads, as demonstrated by the order-of-magnitude difference in run time between the mobile, laptop, and high-end desktop form factors we tested. We also confirmed that outputting arcs instead of lines can lead to a 2$\times$ decrease in execution time (particularly in scenes with high curve content, like \emph{waves} and \emph{mmark}) as predicted by the lower overall primitive count.

@@

Path stroking has received attention in recent years, with \citet{Nehab2020} and \citet{Kilgard2020} both having proposed a complete theory of correct stroking in vector graphics. While there are a number of implementations of path filling on the GPU, the goal of also implementing stroke expansion on the GPU had not yet been realized. Building upon the theory proposed by \citet{Nehab2020}, we implemented an approach to stroke expansion that is GPU-friendly, and achieves high performance on commodity hardware.

@@

\begin{itemize}
    \item The GPU implementation sequentializes some work that could be parallel, and is thus not as well load-balanced as it might be. An appealing future direction is to split the pipeline into separate stages, executed by the GPU as nodes in a work graph[\cite{Patel2024}. This structure is appealing because load balancing is done by hardware.

>>>>

Correctness is another challenging aspect of stroke rendering. We propose that the correctness of stroke outlines be divided into \emph{weak correctness} and \emph{strong correctness.} We define strong correctness as the computation of the outline of a line swept along the segment, maintaining normal orientation, combined with stroke caps and joins. Weak correctness, by contrast, only requires the parallel curves (also known as ``offset curves'') of path segments, combined with caps and the outer contours of joins. Both \cite{Nehab2020} and \cite{Kilgard2020} provide useful definitions of strongly correct stroke rendering, and Nehab in particular describes how to achieve it in stroke expansion. Briefly, when the path curvature exceeds the reciprocal of the stroke half-width, \emph{evolute} segments are required in addition to the parallel curves, as well as \emph{inner contours of joins} when segments are short. The technique described in this paper produces strongly correct outlines according to this definition. Examples of weakly and strongly correct rendering are shown in Figure~\ref{fig:weak_strong}.

@@

The subproblem of approximating parallel curves has spawned extensive literature, none of which is entirely satisfying, especially when it comes to algorithms that can be efficiently evaluated on GPU. An example of a reasonably good algorithm that computes flattened parallel curves is \cite{Yzerman2020}, as used in the Blend2D rendering library. For producing curved outlines, the \cite{Tiller1984} algorithm is commonly implemented and cited, but its performance is quite poor when applied to cubic Béziers. The quadratic Bézier version is adequate, and forms the basis of the algorithm in \cite{Nehab2020}. A variant is also used in Skia~\cite{Skia}; instead of lowering the cubic Bézier to a quadratic approximation and then computing the parallel curve, Skia approximates the parallel curve directly with a quadratic Bézier, then measures the error to determine whether further subdivision is necessary.

Previous work on stroke rendering does not fully solve the problem of fast and correct stroking. Most existing implementations have correctness problems, in particular failing to draw the evolutes and inner joins as required for strong correctness; \cite{Nehab2020} provides a detailed survey. In addition, most implementations are not suitable for running on GPU, requiring sequential execution to produce watertight outlines. One GPU implementation is polar stroking (\cite{Kilgard2020}), a local technique, but among other limitations, it does not bound the Fréchet distance error (rather, it bounds the angle error).

@@

Previous work, particularly \cite{Kilgard2020}, has explained how to break the input path into dash segments of the specified length on GPU using a prefix sum of segment arc lengths. There are implementation tradeoffs to GPU-based dash processing, including potentially a larger number of dispatch stages. We anticipate that the potential performance benefit of implementing dashing on a GPU depends on the scene and other factors in the overall rendering pipeline. While we would expect further performance improvement from GPU-based dashing, we did not fully explore the solution space and we were able to achieve acceptable performance with dash processing on CPU.

@@

There are generally three approaches to such curve approximation. The most straightforward but also least efficient is ``cut then measure,'' usually combined with adaptive subdivision. In this technique, a candidate approximate curve is produced, then the error is measured against the source curve, usually by sampling a number of points along both curves and determining a maximum error (or perhaps some other error norm). If the error is within tolerance, the approximation is accepted. Otherwise, the curve is subdivided (usually at $t = 0.5$) and each half is recursively approximated. A substantial fraction of all curve approximation methods in the literature are of this form, including \cite{Nehab2020}. The main disadvantage is the cost of computing the error metric. Another risk is underestimating the error due to inadequate sampling; this is a particular problem when the source curve contains a cusp.

@@

By far the most efficient approach is an \emph{invertible} error metric. In this approach, the error metric has an analytic inverse, or at least a good numerical approximation. Because the metric is invertible, it can predict the number of subdivisions needed, as well as the parameter value for each subdivision. If the error metric is accurate, then approximation is near-optimal. One example of an invertible error metric is angle step, used in polar stroking (\cite{Kilgard2020}); the number of subdivisions is the total angle subsumed by the curve divided by the angle step size, and the parameter value for each subdivision is the result of solving for a tangent direction. Another widely used invertible error metric is Wang's formula (\cite{Goldman2003}, Section 5.6.3), which gives a bound on the flattening error based on the second derivative of the curve. This metric is conservative but works well in practice; among other applications, it is used in Skia for path flattening. The chief limitation of Wang's method is that it computes a subdivision bound for a Bézier curve without accounting for the displacement of the parallel curve due to stroking. When applied naively to generation of parallel curves, it can undershoot substantially, especially near cusps.

@@

It is common to approximate cubic Béziers to some intermediate curve format more conducive to offsetting and flattening. A number of published solutions (\cite{Yzerman2020}, \cite{Nehab2020}) use quadratic Béziers, as it is well suited for computation of parallel curves. Even so, this curve has some disadvantages. For one, it cannot model an inflection point, so the source curve must be subdivided at inflection points.

@@

The parallel curve of the Euler spiral (also known as ``clothoid'') was characterized by \cite{Wieleitner1907} well over a hundred years ago, and has a straightforward closed-form Cesàro representation, curvature as a function of arc length, of the form $\kappa(s) = c_0(s - s_0)^{-1/2} + c_1$. Euler spirals as an intermediate curve representation for computing parallel curves was proposed in \cite{Levien2021}, which also expands on the Wieleitner reference (including an English translation) and Cesàro equation.

@@

The problem of approximating a curve by a sequence of arc segments has extensive literature, but none of the published solutions are quite suitable for our application. The specific problem of approximating an Euler spiral by arcs is considered in \cite{Meek2004} using a ``cut then measure'' adaptive subdivision scheme, but their solution has poor quality; it scales as $O(1/n^2)$, while $O(1/n^3)$ is attainable. The result was improved ``slightly'' by \cite{Narayan2014}. The literature also contains optimal results, namely \cite{Maier2014} and \cite{Nuntawisuttiwong2021}, but at considerable cost; both approaches claim $O(n^2)$ time complexity. The through-line for all these results is that they are solving a harder problem: adopting the constraint that the generated arc sequence is $G^1$ continuous. While desirable for many applications, this constraint is not needed for rendering a stroke outline. Even with this constraint relaxed, the angle discontinuities of an arc approximation are tiny compared to flattening to lines.

@@

Many techniques for stroke expansion described in the literature apply some lowering of cubic Bézier curves to a simpler curve type that is more tractable for evaluating the parallel curve. Computing parallel curves directly on cubic Bézier curve segments is not very tractable. In particular, the widely cited \cite{Tiller1984} algorithm performs well for quadratic Béziers but significantly worse for cubics.

@@

A typical pattern for converting from one curve type to another is \emph{adaptive subdivision.} An approximate curve is found in the parameter space of the target curve family. The error of the approximation is measured. If the error exceeds the specified tolerance, the curve is subdivided (typically at $t = 0.5$), otherwise the approximation is accepted. Subdivisions are also indicated at special points; for example, since quadratic Béziers cannot represent inflection points, and geometric Hermite interpolation is numerically unstable if the input curve is not convex, lowering to quadratic Béziers also requires calculation of inflection points and subdividing there. A good example of this pattern is \cite{Nehab2020}. One advantage of Euler spirals over quadratic Béziers is that they can represent inflection points just fine, so it is not necessary to solve for the inflection points, or do additional subdivision. Avoiding these alleviates the need for conditional logic and case analysis, which makes the algorithm more amenable to GPU execution.

@@

A key step in approximating one curve with another is evaluating the error of the approximation. A common approach (used in \cite{Nehab2020} among others) is to generate the approximate curve, then measure the distance, often by sampling at multiple points on the curve. All this is potentially slow, with the additional risk of underestimating the error due to undersampling.

@@

Given tangent angles relative to the chord, finding the Euler spiral segment that minimizes total curvature variation is a form of geometric Hermite interpolation. There are a number of published solutions to this problem, involving nontrivial numerical solving techniques: gradient descent~\cite{Kimia2003}, bisection~\cite{Walton2009}, or Newton iteration~\cite{Connor2014}. A more direct approach is to approximate the function as a reasonably low-order polynomial in terms of the endpoint angles. Our approach is to use a 7th order polynomial, which is more precise than the 3rd order polynomial proposed in \cite{Reif2021}, at only modest increased cost. The exact coefficients used are presented in Appendix \ref{appendix:gh}.

@@

\begin{figure*}
    \includegraphics[scale=0.15]{stroke_stills}
    \caption{Stroked cubic Béziers exhibiting strong and weak correctness, rendered by our implementation. The contents of the line soup buffer are visualized as green and blue outlines. Segments with a winding direction that points \emph{up} are colored blue while those pointing \emph{down} are colored green. The curves were adapted from the \emph{stills} dataset by \cite{Nehab2020} and the Skia test corpus.}
    \Description{Selection of stroked cubic Béziers exhibiting strong and weak correctness}
    \label{fig:stroke_stills}
\end{figure*}

As recommended in both \cite{Nehab2020} and \cite{Kilgard2020}, the rendering of a cusp with infinite curvature matches that of a near-cusp. The code to detect near-zero derivatives and re-evaluate is a small amount of non-divergent logic, unlike the additional ``regularization'' pass proposed in Section 3.1 of \cite{Nehab2020} to handle special cases.

@@

Even when the input curve is smooth, its parallel curve contains a cusp when the (signed) curvature equals the offset distance (the half-width of the stroke). In published techniques, dealing with these cusps is a nontrivial effort, and involves numerical methods that are not GPU friendly. In particular, detecting locations in the cubic Bézier where the curvature crosses a given quantity is a medium-degree polynomial, and in general requires numerical techniques for root finding. In the approach of \cite{Nehab2020}, the root finder not only requires a hybrid Newton/bisection method, which requires iteration, but is also recursive in that it uses the roots of a polynomial of one degree lower as a subroutine. In general, polynomials up to cubic can be considered GPU-friendly, while finding cusps in cubic Béziers requires a higher degree.

@@

We present two versions of our stroking method: a sequential CPU stroker and the GPU compute shader outlined in Section \ref{section:gpu-impl}. Both versions can generate stroked outlines with line and arc primitives. We focused our evaluation on the number of output primitives and execution time, using the \emph{timings} dataset provided by \cite{Nehab2020}, in addition to our own curve-intensive scenes to stress the GPU implementation. We present our findings in the remainder of this section.

@@

We measured the timing of our CPU implementation, generating both line and arc primitives, against the Nehab and Skia strokers, which both output a mix of straight lines and quadratic Bézier primitives. The results are shown in Figure~\ref{fig:timings}. The time shown is the total for test files from the \emph{timings} dataset of \cite{Nehab2020}, and the tolerance was 0.25 when adjustable as a parameter.

Similarly to \cite{Nehab2020}'s observations, Skia is the fastest stroke expansion algorithm we measured. The speed is attained with some compromises, in particular an imprecise error estimation that can sometimes generate outlines exceeding the given tolerance. We confirmed that this behavior is still present. The Nehab paper proposes a more careful error metric, but a significant fraction of the total computation time is dedicated to evaluating it.

@@

We instrumented our pipeline to obtain fine-grained pipeline stage execution times using the GPU \emph{timestamp query} feature supported by both Metal and Vulkan. This allowed us to collect measurements for the curve flattening and input processing stages in isolation. We ran the compute pipeline on the \emph{timings} SVG files provided by \cite{Nehab2020}. The largest of these SVGs is \emph{waves} (see rendering \emph{(a)} in Figure \ref{fig:renderings}) which contains 42 stroked paths and a total of 13,308 path segments.

@@

\begin{figure}
    \includegraphics[scale=0.69]{nehab_gpu_timings_strong}
    \caption{Mean GPU execution times of the compute shader that performs stroke expansion and flattening when run on the \cite{Nehab2020} \emph{timings} data set. The timings are shown in milliseconds. The hatched bars show the timings for the version that outputs line primitives while the solid bars show that for arcs. Mali-G78 (the lowest-end GPU we tested) converts \emph{waves} (the largest scene in the set) to lines in \textasciitilde3.5 ms. The compute shader executes at least an order of magnitude faster on the other GPUs.}
    \Description{A stacked bar chart of the GPU timings for stroke expansion from \cite{Nehab2020} timings SVGs}
    \label{fig:nehab-gpu-timings}
\end{figure}

@@

Figures~\ref{fig:nehab-gpu-timings} and \ref{fig:test-scenes-gpu-timings} show the execution times for the two data sets. The entire \cite{Nehab2020} timings set adds up to less than 1.5~ms on all GPUs except the mobile unit. Our algorithm is \textasciitilde14$\times$ slower on the Mali-G78 compared to the M1 Max but it is still capable of processing \emph{waves} in 3.48~ms on average. We observed that the performance of the kernel scales well with increasing GPU ability even at very large workloads, as demonstrated by the order-of-magnitude difference in run time between the mobile, laptop, and high-end desktop form factors we tested. We also confirmed that outputting arcs instead of lines can lead to a 2$\times$ decrease in execution time (particularly in scenes with high curve content, like \emph{waves} and \emph{mmark}) as predicted by the lower overall primitive count.

@@

Path stroking has received attention in recent years, with \cite{Nehab2020} and \cite{Kilgard2020} both having proposed a complete theory of correct stroking in vector graphics. While there are a number of implementations of path filling on the GPU, the goal of also implementing stroke expansion on the GPU had not yet been realized. Building upon the theory proposed by \cite{Nehab2020}, we implemented an approach to stroke expansion that is GPU-friendly, and achieves high performance on commodity hardware.

@@

\begin{itemize}
    \item The GPU implementation sequentializes some work that could be parallel, and is thus not as well load-balanced as it might be. An appealing future direction is to split the pipeline into separate stages, executed by the GPU as nodes in a work graph\cite{Patel2024}. This structure is appealing because load balancing is done by hardware.
